Reasoning About a Machine with Local Capabilities

### 2019-11-08 toplas/paper.tex
toplas: disable todonotes
added orcid for LB
Change journal name to TOPLAS
Address most TOPLAS publisher remarks.

% Metadata Information
\acmJournal{TWEB}
\acmVolume{9}
\acmNumber{4}
\acmArticle{39}
\acmYear{2010}
\acmMonth{3}
\copyrightyear{2009}
%\acmArticleSeq{9}
% Copyright
%\setcopyright{acmcopyright}
\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

@@

\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
% domi: IEEEtran.cls sets font size to 10bp, apparently.
% not sure what the difference is with 10pt.
\DeclareMathSizes{10bp}{8}{7}{6}

@@

\author{Lars~Birkedal}
%\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
%  \position{Position1}
  \department{Department of Computer Science}              %% \department is recommended
  \institution{Aarhus University}            %% \institution is required
  \streetaddress{\AA{}bogade 34}
  \city{Aarhus}
  \postcode{8210}
  \country{Denmark}
}
\email{birkedal@cs.au.dk}          %% \email is recommended
% \author{Lau~Skorstengaard\inst{1} \and Dominique~Devriese\inst{2} \and Lars~Birkedal\inst{1}}
% \institute{Aarhus University \\ \email{\{lau,birkedal\}@cs.au.dk}
% \and Vrije Universiteit Brussel \\ \email{dominique.devriese@vub.be}}

@@

  For instance, we have added an introduction to \sectionname~\nameref{sec:logical-relation} that provides informal intuition about how the logical relation machinery comes into play on a capability machine.

@@

\begin{wrapfigure}{R}{.4\textwidth}
  \centering
  \begin{tikzpicture}[main node/.style={}]
    \node[main node] (7) {$\rwlx$};
    \node[main node] (8) [below left of=7] {$\readwritel$};
    \node[main node] (1) [below right of=7] {$\rwx$};
    \node[main node] (2) [below right of=1] {$\exec$};
    \node[main node] (3) [below right of=2] {$\entry$};
    \node[main node] (4) [below left of=1] {$\readwrite$};
    \node[main node] (5) [below right of=4] {$\readonly$};
    \node[main node] (6) [below right of=5] {$\noperm$};

@@

    \path[every node/.style={font=\sffamily\small}]
    (1l) edge (2l);
  \end{tikzpicture}
  \caption{Permission and locality hierarchy.}
  \label{fig:perm-hier}
\end{wrapfigure}

@@

\begin{lemma}
  \label{lem:correctness-g1}
  Let
\[
    \begin{array}{r@{}lr@{}l}
    c_{\var{adv}} & \;\defeq ((\rwx,\glob),\dots) & c_{g1} & \;\defeq ((\entry,\glob),\dots)
    \end{array}
\]
  and otherwise make assumptions about capabilities and linking similar to Lemma~\ref{lem:correctness-f1}.
  Then if $
  (\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step^* (\halted,m'),$
  then $m'(\flag) = 0$.
\end{lemma}
\begin{toplas}
  \begin{proof}[Proof sketch]
    Define a world $W_1$ with the following regions
    \begin{itemize}
    \item A malloc region, $\iota_{\malloc,0}$.
    \item A permanent region for the linking table that only accepts $\ms_\link$ and requires everything to be in $\stdvr$.
    \item A stack region, $\iota^{\pwl}_{b_\stk,e_\stk}$.
    \item An adversary region, $\iota^{\nwl,p}_{b_\adv,e_\adv}$
    \item A permanent static region for the flag and \texttt{g1}, i.e.\ a region that only
      accepts $\ms_\flag$ and $\ms_{g1}$.
    \end{itemize}
    If we can show
    \begin{equation}
      (\reg_0[\pcreg \mapsto c_\adv,r_\stk\mapsto c_\stk,r_1\mapsto c_{g1}],\ms_\malloc \uplus \ms_\link \uplus \ms_\stk \uplus \ms_\adv\uplus \ms_\flag  \uplus \ms_{g1}) \in \observations(W_1),\label{eq:g1:main-O-res}
  \end{equation}
  \footnote{We ignore step indexes in this proof sketch.}then the $\observations$-relation ensures that a successfully halting configuration terminates in a memory that respects a private future world of $W$ which in particular means that it respects the permanent static region that governs the assertion flag.

>>>>

% Metadata Information
\acmJournal{TOPLAS}
\acmVolume{0}
\acmNumber{0}
\acmArticle{0}
\acmYear{2019}
\acmMonth{0}
\copyrightyear{2019}
%\acmArticleSeq{9}
% Copyright
%\setcopyright{acmcopyright}
\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

@@

% \usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
\usepackage[disable]{todonotes}
% domi: IEEEtran.cls sets font size to 10bp, apparently.
% not sure what the difference is with 10pt.
\DeclareMathSizes{10bp}{8}{7}{6}

@@

\author{Lars~Birkedal}
%\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{0000-0003-1320-0098}          %% \orcid is optional
\affiliation{
%  \position{Position1}
  \department{Department of Computer Science}              %% \department is recommended
  \institution{Aarhus University}            %% \institution is required
  \streetaddress{\AA{}bogade 34}
  \city{Aarhus}
  \postcode{8210}
  \country{Denmark}
}
\email{birkedal@cs.au.dk}          %% \email is recommended
% \author{Lau~Skorstengaard\inst{1} \and Dominique~Devriese\inst{2} \and Lars~Birkedal\inst{1}}
% \institute{Aarhus University \\ \email{\{lau,birkedal\}@cs.au.dk}
% \and Vrije Universiteit Brussel \\ \email{dominique.devriese@vub.be}}

@@

  For instance, we have added an introduction to \sectionname~\ref{sec:logical-relation} that provides informal intuition about how the logical relation machinery comes into play on a capability machine.

@@

\begin{figure}
  \centering
  \begin{tikzpicture}[main node/.style={}]
    \node[main node] (7) {$\rwlx$};
    \node[main node] (8) [below left of=7] {$\readwritel$};
    \node[main node] (1) [below right of=7] {$\rwx$};
    \node[main node] (2) [below right of=1] {$\exec$};
    \node[main node] (3) [below right of=2] {$\entry$};
    \node[main node] (4) [below left of=1] {$\readwrite$};
    \node[main node] (5) [below right of=4] {$\readonly$};
    \node[main node] (6) [below right of=5] {$\noperm$};

@@

    \path[every node/.style={font=\sffamily\small}]
    (1l) edge (2l);
  \end{tikzpicture}
  \caption{Permission and locality hierarchy.}
  \label{fig:perm-hier}
\end{figure}

@@

\begin{lemma}
  \label{lem:correctness-g1}
  Let
\[
    \begin{array}{r@{}lr@{}l}
    c_{\var{adv}} & \;\defeq ((\rwx,\glob),\dots) & c_{g1} & \;\defeq ((\entry,\glob),\dots)
    \end{array}
\]
  and otherwise make assumptions about capabilities and linking similar to Lemma~\ref{lem:correctness-f1}.
  Then if $
  (\reg_0\update{\pcreg}{c_\adv}\update{r_\stk}{c_\stk}\update{r_1}{c_{g1}},m) \step^* (\halted,m'),$
  then $m'(\flag) = 0$.
\end{lemma}
\begin{toplas}
  \begin{proof}[Proof sketch]
    Define a world $W_1$ with the following regions
    \begin{itemize}
    \item A malloc region, $\iota_{\malloc,0}$.
    \item A permanent region for the linking table that only accepts $\ms_\link$ and requires everything to be in $\stdvr$.
    \item A stack region, $\iota^{\pwl}_{b_\stk,e_\stk}$.
    \item An adversary region, $\iota^{\nwl,p}_{b_\adv,e_\adv}$
    \item A permanent static region for the flag and \texttt{g1}, i.e.\ a region that only
      accepts $\ms_\flag$ and $\ms_{g1}$.
    \end{itemize}
    If we can show\footnote{We ignore step indexes in this proof sketch.}
    \begin{equation}
      (\reg_0[\pcreg \mapsto c_\adv,r_\stk\mapsto c_\stk,r_1\mapsto c_{g1}],\ms_\malloc \uplus \ms_\link \uplus \ms_\stk \uplus \ms_\adv\uplus \ms_\flag  \uplus \ms_{g1}) \in \observations(W_1),\label{eq:g1:main-O-res}
  \end{equation}
  then the $\observations$-relation ensures that a successfully halting configuration terminates in a memory that respects a private future world of $W$ which in particular means that it respects the permanent static region that governs the assertion flag.
